RIScatter: Unifying Backscatter Communication and Reconfigurable Intelligent Surface

### 2023-03-01 arxiv.tex
:triangular_flag_on_post: JSAC 230301

% algpseudocode
\makeatletter
\renewcommand{\fnum@algorithm}{\fname@algorithm{} \thealgorithm:}
\makeatother
\algrenewcommand{\algorithmicrequire}{\textbf{Input:}}
\algrenewcommand{\algorithmicensure}{\textbf{Output:}}
\algrenewcommand{\algorithmicwhile}{\textbf{While}}
\algrenewcommand{\algorithmicend}{\textbf{End}}
\algrenewcommand{\algorithmicrepeat}{\textbf{Repeat}}
\algrenewcommand{\algorithmicuntil}{\textbf{Until}}
\algrenewcommand{\algorithmicdo}{}

@@

\begin{IEEEkeywords}
	Input distribution design, energy detection, \gls{sic}-free receiver, symbol-level precoding, active-passive coexisting network, backscatter communication, reconfigurable intelligent surface, ambient backscatter communication, symbiotic radio.
\end{IEEEkeywords}

@@

	% \begin{subsection}{Related Works}
	As a special case of \gls{cr}, active and passive transmissions coexist and interplay in \gls{ambc} and \gls{sr}.
	Such a coexistence is classified into commensal (overlay), parasitic (underlay), and competitive paradigms, and their achievable rate and outage performance were investigated in \cite{Guo2019b,Ding2020}.
	For the co-located cooperative receiver, the \gls{ber} performance of \gls{ml}, linear, and \gls{sic} detectors are derived over flat fading channels \cite{Yang2018}.
	However, the work assumed equal symbol duration and perfect synchronization for primary and backscatter links.
	Importantly, active-passive coexisting networks have three special and important properties:
	\begin{enumerate}
		\item Primary and backscatter symbols are superimposed by \emph{double modulation} (i.e., multiplication coding);
		\item Backscatter signal strength is much weaker than primary due to the \emph{double fading} effect;
		\item The spreading factor (i.e., backscatter symbol duration over primary) is usually large\footnote{The load-switching interval of low-power backscatter modulators is usually \num{0.1} to \qty{10}{\us} \cite{Torres2021}, accounting for a typical spreading factor between \num{10} and \num{e3}.}.
	\end{enumerate}
	The second property motivated \cite{Long2020a,Liang2020,Guo2019b,Ding2020,Zhou2019a,Wu2021a,Xu2021a,Yang2021a,Yang2018,Han2021,Zhang2022} to view \gls{sr} as a multiplicative \gls{noma} and perform \gls{sic} from primary to backscatter link.
	During primary decoding, the backscatter signal can be modelled as channel uncertainty or multiplicative interference when the spreading factor is large or small, respectively.
	Decoding each backscatter symbol also requires multiple \gls{sic} followed by a \gls{mrc} and is operation-intensive and \gls{csi}-sensitive.
	% Such a \gls{sic} receiver is operation-intensive and \gls{csi}-sensitive because it involves $N$ re-encoding, precoding, subtraction together with a time-domain \gls{mrc} to decode each backscatter symbol.
	When the spreading factor is sufficiently large, the primary achievable rate under semi-coherent detection\footnote{In this paper, semi-coherent detection refers to the primary/backscatter decoding with known \gls{csi} and unknown backscatter/primary symbols.} asymptotically approaches its coherent counterpart and both links are decoded interference-free \cite{Long2020a}.
	% However, it significantly increases \gls{sic} operations per backscatter symbol and severely limits the backscatter \emph{throughput}.
	However, it severely limits the backscatter throughput and requires numerous \gls{sic} per backscatter symbol.

@@

	On the other hand, static \gls{ris} design with fixed reflection pattern per channel block has been extensively studied in wireless communication, sensing, and power literature \cite{Wu2018,Zhang2019a,Lin2022,Liu2022,Feng2022,Zhao2022}.
	Dynamic \gls{ris} performs time sharing between different phase shifts and introduces artificial channel diversity within each channel block.
	It was first proposed to fine-tune the \gls{ofdm} resource blocks \cite{Yang2020} then extended to the downlink power and uplink information phases of \gls{wpcn} \cite{Wu2021,Wu2021d,Hua2022a}.
	However, dynamic \gls{ris} carries no information since the reflection state at each time slot is known to the receiver.
	\gls{ris} can also be used as an index-modulated information source and prototypes have been developed for \gls{psk} \cite{Tang2019a} and \gls{qam} \cite{Dai2020a}.
	From an information-theoretic perspective, the authors of \cite{Karasik2020} reported that joint transmitter-\gls{ris} encoding achieves the capacity of \gls{ris}-aided finite-input channel and using \gls{ris} as a naive passive beamformer to maximize the receive \gls{snr} is generally suboptimal.
	This inspired \gls{ris}-empowered \gls{bc} \cite{Liu2019d,Bereyhi2020,Xu2020b,Zhang2021d,Hu2021b,Hua2022,Basar2020,Ma2020a,Yuan2021,Hu2021a} to combine passive beamforming and backscatter modulation in the overall reflection pattern.
	In particular, \emph{symbol level precoding} maps the information symbols to the optimized \gls{ris} coefficient sets \cite{Liu2019d,Bereyhi2020}, \emph{overlay modulation} superposes the information symbols over a common auxiliary matrix \cite{Xu2020b,Zhang2021d,Hu2021b,Hua2022}, \emph{spatial modulation} switches between the reflection coefficient sets that maximize \gls{snr} at different receive antennas \cite{Basar2020,Ma2020a,Yuan2021}, and \emph{index modulation} employs dedicated reflection elements (resp. information elements) for passive beamforming (resp. backscatter modulation) \cite{Hu2021a}.
	However, those joint designs incur advanced hardware architecture and high optimization complexity.
	Relevant literature also considers either Gaussian codebook \cite{Guo2019b,Ding2020,Long2020a,Zhou2019a,Wu2021a,Xu2021a,Yang2021a,Hu2021b} that is impractical for low-power nodes or finite equiprobable inputs \cite{Yang2018,Liang2020,Han2021,Zhang2022,Liu2019d,Bereyhi2020,Xu2020b,Zhang2021d,Hua2022,Basar2020,Ma2020a,Yuan2021,Hu2021a} that does not fully exploit the \gls{csi} and properties of active-passive coexisting networks.
	% \end{subsection}
	Those problems are addressed in this paper and the contributions are summarized below.

@@

	\emph{Second,} we address the first challenge and propose a practical receiver that semi-coherently decodes the backscatter information, recovers the reflection pattern, and coherently decodes the primary link.
	We consider backscatter energy detection and formulate a \gls{dmmac} over disjoint decision regions.
	Thanks to the double modulation and symbol-level passive precoding, the semi-coherent backscatter decoding also acts as part of channel training.
	Once the backscatter information is successfully decoded, its contribution can be modelled within composite channel as dynamic passive beamforming rather than re-precoded and cancelled.
	It requires only one energy comparison and re-encoding per backscatter symbol and is much efficient than conventional decoding based on \gls{sic} and \gls{mrc}.
	It is also suitable for arbitrary input distribution and spreading factor, which significantly improves backscatter throughput.
	This is the first paper to consider practical backscatter-primary decoding scheme to exploit dynamic passive beamforming at the co-located receiver.

@@

\begin{section}{RIScatter}
	\begin{subsection}{Principles}
		\gls{rf} wave scattering/reflecting are often manipulated by passive antennas or programmable metamaterial \cite{Liang2022}.
		% can be described by a unified signal model and implemented by passive antennas or programmable metamaterial \cite{Liang2022}.
		% An antenna-based scatterer usually consists of an integrated antenna, a load-switching modulator, an energy harvester, and on-chip components (e.g., microcontroller and sensors) \cite{Dobkin2012}.
		% An antenna-based scatterer usually consists of an integrated antenna, a load-switching modulator, an energy harvester, and on-chip components (e.g., microcontroller and sensors) \cite{Dobkin2012}.
		An antenna scatterer first receives the impinging signals then reradiates some back to the space and dissipates the remaining \cite{Dobkin2012}.
		% In comparison, a typical metamaterial-based scatterer comprises an outer metamaterial layer of numerous sub-wavelength metallic/dielectric patches (with tunable permittivity/permeability), a middle copper plate layer that avoids leakage, an inner circuit board layer that adjusts patch responses, and an integrated microcontroller/\glsentryshort{fpga} for coordination and control \cite{Wu2020}.
		% In comparison, a typical metamaterial-based scatterer comprises an outer metamaterial layer of numerous sub-wavelength metallic/dielectric patches (with tunable permittivity/permeability), a middle copper plate layer that avoids leakage, an inner circuit board layer that adjusts patch responses, and an integrated microcontroller/\glsentryshort{fpga} for coordination and control \cite{Wu2020}.
		An ideal metamaterial scatterer reflects the wave at the space-metamaterial boundary without receiving them and mainly applies a phase shift on top.
		% In practice, both types of scatterers have {finite} reflection states with non-zero reflection loss.
		% In general, the scattered signal consists of a structural mode component that consistently contributes to environment multipath and can be modelled within \gls{csi}, and an \emph{antenna} mode component that depends on the impedance mismatch and can be used for backscatter modulation and/or passive beamforming \cite{Hansen1989}.
		In general, the scattered signal consists of a structural mode component that consistently contributes to environment multipath and an antenna mode component that depends on the impedance mismatch \cite{Hansen1989}.
		The former is usually modelled within \gls{csi} while the latter is mainly used for backscatter modulation and/or passive beamforming.
		For an antenna (resp. metamaterial) scatterer with $M$ reflection states, the reflection coefficient at state $m \in \mathcal{M} \triangleq \{1,\ldots,M\}$ is
			\Gamma_m = \frac{Z_m - Z^*}{Z_m + Z},
			\label{eq:reflection_coefficient}
		where $Z_m$ is the antenna load (resp. metamaterial unit) impedance at state $m$ and $Z$ is the antenna input (resp. medium characteristic) impedance.
		\gls{bc} treats scatterers as information sources that \emph{randomly} choosing the reflection state based on information symbol.
		For $M$-ary \gls{qam}, constellation point $c_m$ maps to reflection coefficient $\Gamma_m$ by \cite{Thomas2012a}
			\Gamma_m = \alpha \frac{c_m}{\max_{m'} \lvert c_{m'} \rvert},
			\label{eq:backscatter_modulation}
		where $\alpha \in \mathbb{I}$ is the common amplitude scattering ratio at the direction of interest.
		In contrast, \gls{ris} treats scatterers as channel adaptors that \emph{deterministically} choosing the reflection state based on \gls{csi}.
		For a \gls{ris} element with $M$ available states, phase shift $\theta_m$ maps to reflection coefficient $\Gamma_m$ by \cite{Wu2018}
			\Gamma_m = \beta_m \exp(j \theta_m),
			\label{eq:passive_beamforming}
		where $\beta_m \in \mathbb{I}$ is the overall amplitude scattering ratio at state $m$.\footnote{Most existing \gls{ris} literature assumes lossless reflection $\beta_m=1$, $\forall m$.}
		% A natural question is, is it possible to use a scatterer
		For a scatter node with given load impedance set, \gls{bc} employs $\Gamma_m$ only as an information codeword while \gls{ris} employs $\Gamma_m$ only as a passive beamforming codeword.
		RIScatter generalizes both from a probabilistic perspective where each node \emph{semi-randomly} chooses the reflection state with the guidance of probability $P(\Gamma_m)$.
		The input distribution is flexibly designed as a joint function of the information source, \gls{csi}, and \gls{qos}.
		\gls{bc} and (static) \gls{ris} can be regarded as its extreme cases where the input distribution boils down to uniform and degenerate, respectively.
		Fig.~\ref{fg:scatter_comparison} compares the input distribution and reflection pattern of RIScatter to \gls{bc} and \gls{ris}.
			\subfloat[Input Distribution]{
				\resizebox{0.8\columnwidth}{!}{
					\input{assets/illustration/input_distribution.tex}
				\label{fg:input_distribution}
			\\
			\subfloat[Reflection Pattern]{
				\resizebox{0.8\columnwidth}{!}{
					\input{assets/illustration/time_structure.tex}
				\label{fg:time_structure}
				Input distribution and reflection pattern of scattering applications.
				``PB'', ``BB'', and ``CB'' refer to primary symbol block, backscatter symbol block, and channel block, respectively.
				Shadowing means presence of primary link.
				In this example, the optimal passive beamformer is state \num{2} and its input probability for RIScatter is \num{0.75}.
				% RIScatter node transmit symbol \num{1}
				The spreading factor is \num{4} for RIScatter and \num{8} for \gls{ambc}/\gls{sr}.
			\label{fg:scatter_comparison}
			RIScatter is different from conventional dynamic passive beamforming since the reflection pattern conveys additional information.
			% There are many ways to order
			% The latter only involves time sharing between different states but there are many ways to ensure within a channel block.
			The latter only involves time sharing between different states and the actual order has no impact on the (ergodic) primary performance.
			For RIScatter, different state combinations correspond to different messages and the impact reduces to dynamic passive beamforming upon successful backscatter decoding.
			% However, different combinations correspond to different backscatter messages and RIScatter impact is determined dynamic passive beamforming once successfully decoded.

@@

			\resizebox{0.8\columnwidth}{!}{
				\input{assets/illustration/energy_distribution.tex}

@@

	Problem \eqref{op:weighted_sum_rate} generalizes conventional \gls{bc} by allowing \gls{csi}- and \gls{qos}-adaptive input distribution and decision region design.
	It also generalizes the discrete \gls{ris} phase shift selection by allowing stochastic reflection, relaxing the feasible domain from the vertices of $M$-dimensional probability simplex to the simplex itself.
	Problem \eqref{op:weighted_sum_rate} is highly non-convex and we propose a \gls{bcd} algorithm that iteratively updates $\{\boldsymbol{p}_k\}_{k \in \mathcal{K}}$, $\boldsymbol{w}$ and $\boldsymbol{t}$.

@@

		\begin{remark}
			Energy detector design has no impact on the primary achievable rate since the composite channel \eqref{eq:equivalent_channel} can always be reconstructed by re-encoding.
			It suggests any thresholding scheme that maximize the total backscatter rate \eqref{eq:backscatter_mutual_information} is also optimal for problem \eqref{op:decision_threshold}.
			\label{re:backscatter_decision}
		\end{remark}

@@

\begin{section}{Simulation Results}
	In this section, we provide numerical results to evaluate the proposed input distribution, active beamforming, and backscatter decision designs for the considered RIScatter system.
	We assume the AP-user distance is \qty{10}{\meter} and at least one RIScatter nodes are randomly dropped in a disk centered at the user with radius $r$.
	% The \gls{ap} is with an average transmit power budget $P=\qty{36}{\dBm}$, all nodes employs $M$-\gls{qam} with amplitude scattering ratio $\alpha=0.5$, and the user is with average noise power $\sigma_v^2=\qty{-40}{\dBm}$.
	The \gls{ap} is with an average transmit power budget $P=\qty{36}{\dBm}$ and all nodes employs $M$-\gls{qam} with $\alpha=0.5$.
	For all channels involved, we consider a distance-dependent path loss model
	\begin{equation}
		L(d) = L_0 \biggl(\frac{d_0}{d}\biggr)^\gamma,
	\end{equation}
	together with a Rician fading model
	\begin{equation}
		\boldsymbol{H} = \sqrt{\frac{\kappa}{1+\kappa}} \bar{\boldsymbol{H}} + \sqrt{\frac{1}{1+\kappa}} \tilde{\boldsymbol{H}},
	\end{equation}
	where $d$ is the transmission distance, $L_0=-\qty{30}{\dB}$ is the reference path loss at $d_0=\qty{1}{\meter}$, $\kappa$ is the Rician K-factor, $\bar{\boldsymbol{H}}$ is the deterministic line-of-sight component with unit-magnitude entries, and $\tilde{\boldsymbol{H}}$ is the Rayleigh fading component with standard \gls{iid} \gls{cscg} entries.
	We choose $\gamma_{\text{D}}=2.6$, $\gamma_{\text{F}}=2.4$, $\gamma_{\text{B}}=2$, and $\kappa_{\text{D}}=\kappa_{\text{F}}=\kappa_{\text{B}}=5$ for direct, forward and backward links.
	The finite decision threshold domain $\mathcal{T}$ is obtained by $b$-bit uniform discretization over the critical interval defined by the confidence bounds of edge hypotheses (i.e., lower bound of $\mathcal{H}_1$ and upper bound of $\mathcal{H}_L$) with confidence $1-\varepsilon$.
	We choose $b=9$ and $\varepsilon=\num{e-3}$.
	All achievable rate points/regions are averaged over \num{e3} channel realizations.

@@

		\begin{subsubsection}{Convergence}
			\begin{figure}[!t]
				\centering
				\resizebox{0.75\columnwidth}{!}{
					\input{assets/simulation/wsr_convergence.tex}
				}
				\caption{Typical convergence curves at $\rho=0$ for $Q=4$, $K=8$, $M=2$, $N=20$, $\sigma_v^2=\qty{-40}{\dBm}$ and $r=\qty{2}{\meter}$.}
				\label{fg:wsr_convergence}
			\end{figure}

			In Fig.~\ref{fg:wsr_convergence}, we plot the weighted sum of primary and total backscatter rates at $\rho=0$ for \gls{kkt}, \gls{pgd} and \gls{bcd} algorithms on the first call.
			For $K=8$ and $M=2$, Algorithm \ref{al:input_distribution} typically takes around \num{100} fast iterations to converge to the \gls{kkt} input distribution.
			For $Q=4$, around \num{10} iterations are required for Algorithm \ref{al:active_beamforming} to converge.
			Overall, the initial \gls{bcd} algorithm requires at most \num{5} iterations to converge.
			At the following points (not presented here), the convergence of all three algorithms are much faster thanks to the progressive initialization.
			We conclude the proposed algorithms have good convergence performances.
		\end{subsubsection}
	\end{subsection}

@@

			\caption{Typical achievable rate region/points of scattering applications for $Q=1$, $K=1$, $M=4$, $N=e3$, $\sigma_v^2=\qty{-40}{\dBm}$ and $r=\qty{2}{\meter}$.}

>>>>

% algpseudocode
\makeatletter
\renewcommand{\fnum@algorithm}{\fname@algorithm{} \thealgorithm:}
\makeatother

\algrenewcommand{\algorithmicrequire}{\textbf{Input:}}
\algrenewcommand{\algorithmicensure}{\textbf{Output:}}
\algrenewcommand{\algorithmicwhile}{\textbf{While}}
\algrenewcommand{\algorithmicend}{\textbf{End}}
\algrenewcommand{\algorithmicrepeat}{\textbf{Repeat}}
\algrenewcommand{\algorithmicuntil}{\textbf{Until}}
\algrenewcommand{\algorithmicdo}{}

@@

\begin{IEEEkeywords}
	Input distribution design, symbol-level precoding, active-passive coexisting network, \gls{sic}-free receiver, energy detection, backscatter communication, reconfigurable intelligent surface, ambient backscatter communication, symbiotic radio.
\end{IEEEkeywords}

@@

	% \begin{subsection}{Related Works}
	As a special case of \gls{cr}, active and passive transmissions coexist and interplay in \gls{ambc} and \gls{sr}.
	Such a coexistence is classified into commensal (overlay), parasitic (underlay), and competitive (interfering) paradigms, and their achievable rate and outage performance were investigated in \cite{Guo2019b,Ding2020}.
	For the co-located cooperative receiver, the \gls{ber} performance of \gls{ml}, linear, and \gls{sic} detectors are derived over flat fading channels \cite{Yang2018}.
	However, the work assumed equal symbol duration and perfect synchronization for primary and backscatter links.
	Importantly, active-passive coexisting networks have three special and important properties:
	\begin{enumerate}
		\item Primary and backscatter symbols are superimposed by \emph{double modulation} (i.e., multiplication coding);
		\item Backscatter signal strength is much weaker than primary due to the \emph{double fading} effect;
		\item The spreading factor (i.e., backscatter symbol duration over primary) is usually large\footnote{The load-switching interval of low-power backscatter modulators is usually {0.1} to \SI{10}{\us} \cite{Torres2021}, accounting for a typical spreading factor between $10$ and $10^3$.}.
	\end{enumerate}
	The second property motivated \cite{Long2020a,Liang2020,Guo2019b,Ding2020,Zhou2019a,Wu2021a,Xu2021a,Yang2021a,Yang2018,Han2021,Zhang2022} to view \gls{sr} as a multiplicative \gls{noma} and perform \gls{sic} from primary to backscatter link.
	During primary decoding, the backscatter signal can be modelled as channel uncertainty or multiplicative interference when the spreading factor is large or small, respectively.
	Decoding each backscatter symbol also requires multiple \gls{sic} followed by a \gls{mrc} and is operation-intensive and \gls{csi}-sensitive.
	% Such a \gls{sic} receiver is operation-intensive and \gls{csi}-sensitive because it involves $N$ re-encoding, precoding, subtraction together with a time-domain \gls{mrc} to decode each backscatter symbol.
	When the spreading factor is sufficiently large, the primary achievable rate under semi-coherent detection\footnote{In this paper, semi-coherent detection refers to the primary/backscatter decoding with known \gls{csi} and unknown backscatter/primary symbols.} asymptotically approaches its coherent counterpart and both links are decoded interference-free \cite{Long2020a}.
	% However, it significantly increases \gls{sic} operations per backscatter symbol and severely limits the backscatter \emph{throughput}.
	However, this severely limits the backscatter throughput and requires numerous \gls{sic} per backscatter symbol.

@@

	On the other hand, static \gls{ris} design with fixed reflection pattern per channel block has been extensively studied in wireless communication, sensing, and power literature \cite{Wu2018,Zhang2019a,Lin2022,Liu2022,Feng2022,Zhao2022}.
	Dynamic \gls{ris} performs time sharing between different phase shifts and introduces artificial channel diversity within each channel block.
	It was first proposed to fine-tune the \gls{ofdm} resource blocks \cite{Yang2020} then extended to the downlink power and uplink information phases of \gls{wpcn} \cite{Wu2021,Wu2021d,Hua2022a}.
	However, dynamic \gls{ris} carries \emph{no} additional information since the reflection state at each time slot is known to the receiver.
	\gls{ris} can also be used as an information source and prototypes have been developed for \gls{psk} \cite{Tang2019a} and \gls{qam} \cite{Dai2020a}.
	From an information-theoretic perspective, the authors of \cite{Karasik2020} reported that joint transmitter-\gls{ris} encoding achieves the capacity of \gls{ris}-aided finite-input channel and using \gls{ris} as a naive passive beamformer to maximize the receive \gls{snr} is generally suboptimal.
	This inspired \gls{ris}-empowered \gls{bc} \cite{Liu2019d,Bereyhi2020,Xu2020b,Zhang2021d,Hu2021b,Hua2022,Basar2020,Ma2020a,Yuan2021,Hu2021a} to combine passive beamforming and backscatter modulation in the overall reflection pattern.
	In particular, \emph{symbol level precoding} maps the information symbols to the optimized \gls{ris} coefficient sets \cite{Liu2019d,Bereyhi2020}, \emph{overlay modulation} superposes the information symbols over a common auxiliary matrix \cite{Xu2020b,Zhang2021d,Hu2021b,Hua2022}, \emph{spatial modulation} switches between the reflection coefficient sets that maximize \gls{snr} at different receive antennas \cite{Basar2020,Ma2020a,Yuan2021}, and \emph{index modulation} employs dedicated reflection elements (resp. information elements) for passive beamforming (resp. backscatter modulation) \cite{Hu2021a}.
	However, those joint designs incur advanced hardware architecture and high optimization complexity.
	Relevant literature also considers either Gaussian codebook \cite{Guo2019b,Ding2020,Long2020a,Zhou2019a,Wu2021a,Xu2021a,Yang2021a,Hu2021b} that is impractical for low-power nodes or finite equiprobable inputs \cite{Yang2018,Liang2020,Han2021,Zhang2022,Liu2019d,Bereyhi2020,Xu2020b,Zhang2021d,Hua2022,Basar2020,Ma2020a,Yuan2021,Hu2021a} that does not fully exploit the \gls{csi} and properties of active-passive coexisting networks.
	% \end{subsection}
	Those problems are addressed in this paper and the contributions are summarized below.

@@

	\emph{Second,} we address the first challenge and propose a practical receiver that semi-coherently decodes the backscatter information, recovers the reflection pattern, and coherently decodes the primary link.
	We consider backscatter energy detection and formulate a \gls{dmmac} over disjoint decision regions.
	Thanks to the double modulation and symbol-level passive precoding, the semi-coherent backscatter decoding is essentially part of channel training.
	Once the backscatter information is successfully decoded, its contribution can be modelled within composite channel as dynamic passive beamforming (rather than re-precoded and cancelled).
	It requires only one energy comparison and re-encoding per backscatter symbol and is much efficient than conventional decoding based on \gls{sic} and \gls{mrc}.
	It is also suitable for arbitrary input distribution and spreading factor, which significantly improves backscatter throughput.
	This is the first paper to consider practical backscatter-primary decoding scheme to exploit backscatter modulation as dynamic passive beamforming at the co-located receiver.

@@

\begin{section}{RIScatter}
	\begin{subsection}{Principles}
		\gls{rf} wave scattering/reflecting are often manipulated by passive antennas or programmable metamaterial \cite{Liang2022}.
		% can be described by a unified signal model and implemented by passive antennas or programmable metamaterial \cite{Liang2022}.
		% An antenna-based scatterer usually consists of an integrated antenna, a load-switching modulator, an energy harvester, and on-chip components (e.g., microcontroller and sensors) \cite{Dobkin2012}.
		% An antenna-based scatterer usually consists of an integrated antenna, a load-switching modulator, an energy harvester, and on-chip components (e.g., microcontroller and sensors) \cite{Dobkin2012}.
		The former first receives the impinging signals then reradiates some back to the space and dissipates the remaining \cite{Dobkin2012}.
		% In comparison, a typical metamaterial-based scatterer comprises an outer metamaterial layer of numerous sub-wavelength metallic/dielectric patches (with tunable permittivity/permeability), a middle copper plate layer that avoids leakage, an inner circuit board layer that adjusts patch responses, and an integrated microcontroller/\glsentryshort{fpga} for coordination and control \cite{Wu2020}.
		% In comparison, a typical metamaterial-based scatterer comprises an outer metamaterial layer of numerous sub-wavelength metallic/dielectric patches (with tunable permittivity/permeability), a middle copper plate layer that avoids leakage, an inner circuit board layer that adjusts patch responses, and an integrated microcontroller/\glsentryshort{fpga} for coordination and control \cite{Wu2020}.
		In contrast, the latter reflects the wave at the space-metamaterial boundary without receiving them and mainly applies a phase shift.
		% In practice, both types of scatterers have {finite} reflection states with non-zero reflection loss.
		% In general, the scattered signal consists of a structural mode component that consistently contributes to environment multipath and can be modelled within \gls{csi}, and an \emph{antenna} mode component that depends on the impedance mismatch and can be used for backscatter modulation and/or passive beamforming \cite{Hansen1989}.
		The scattered signal consists of a structural mode component that consistently contributes to environment multipath (usually modelled within \gls{csi}) and an antenna mode component that depends on the impedance mismatch (mainly used for backscatter modulation and/or passive beamforming) \cite{Hansen1989}.
		% The former is usually modelled within \gls{csi} while the latter is mainly used for backscatter modulation and/or passive beamforming.
		For an antenna (resp. metamaterial) scatterer with $M$ reflection states, the reflection coefficient at state $m \in \mathcal{M} \triangleq \{1,\ldots,M\}$ is
		\begin{equation}
			\Gamma_m = \frac{Z_m - Z^*}{Z_m + Z},
			\label{eq:reflection_coefficient}
		\end{equation}
		where $Z_m$ is the antenna load (resp. metamaterial unit) impedance at state $m$ and $Z$ is the antenna input (resp. medium characteristic) impedance.

		\gls{bc} treats scatterers as information sources that \emph{randomly} choosing the reflection state based on information symbol.
		For $M$-ary \gls{qam}, constellation point $c_m$ maps to reflection coefficient $\Gamma_m$ by \cite{Thomas2012a}
		\begin{equation}
			\Gamma_m = \alpha \frac{c_m}{\max_{m'} \lvert c_{m'} \rvert},
			\label{eq:backscatter_modulation}
		\end{equation}
		where $\alpha \in \mathbb{I}$ is the common amplitude scattering ratio at the direction of interest.
		In contrast, \gls{ris} treats scatterers as channel adaptors that \emph{deterministically} choosing the reflection state based on \gls{csi}.
		For a \gls{ris} element with $M$ available states, phase shift $\theta_m$ maps to reflection coefficient $\Gamma_m$ by \cite{Wu2018}
		\begin{equation}
			\Gamma_m = \beta_m \exp(j \theta_m),
			\label{eq:passive_beamforming}
		\end{equation}
		where $\beta_m \in \mathbb{I}$ is the overall amplitude scattering ratio at state $m$.\footnote{Most existing \gls{ris} literature assumes lossless reflection $\beta_m=1$, $\forall m$.}
		% A natural question is, is it possible to use a scatterer
		For a scatter node with given load impedance set, \gls{bc} employs $\Gamma_m$ only as an information codeword while \gls{ris} employs $\Gamma_m$ only as a passive beamforming codeword.
		RIScatter generalizes both from a probabilistic perspective where each node \emph{semi-randomly} chooses the reflection state with the guidance of probability $P(\Gamma_m)$.
		The input distribution is flexibly designed as a joint function of the information source, \gls{csi}, and \gls{qos} of both links.
		\gls{bc} and (static) \gls{ris} can be regarded as its extreme cases where the input distribution boils down to uniform and degenerate, respectively.
		Fig.~\ref{fg:scatter_comparison} compares the input distribution and reflection pattern of RIScatter to \gls{bc} and \gls{ris}.
		\begin{figure}[!t]
			\centering
			\subfloat[Input Distribution]{
				\resizebox{0.8\columnwidth}{!}{
					\input{assets/illustration/input_distribution.tex}
				}
				\label{fg:input_distribution}
			}
			\\
			\subfloat[Reflection Pattern]{
				\resizebox{0.8\columnwidth}{!}{
					\input{assets/illustration/time_structure.tex}
				}
				\label{fg:time_structure}
			}
			\caption{
				Input distribution and reflection pattern of scattering applications.
				``PB'', ``BB'', and ``CB'' refer to primary symbol block, backscatter symbol block, and channel block, respectively.
				Shadowing means presence of primary link.
				In this example, the optimal passive beamformer is state {2} and its input probability for RIScatter is {0.75}.
				% RIScatter node transmit symbol {1}
				The spreading factor is {4} for RIScatter and {8} for \gls{ambc}/\gls{sr}.
			}
			\label{fg:scatter_comparison}
		\end{figure}
		\begin{remark}
			RIScatter is different from conventional dynamic passive beamforming since the reflection pattern conveys additional information.
			% There are many ways to order
			% The latter only involves time sharing between different states but there are many ways to ensure within a channel block.
			The latter only involves time sharing between different states and the actual order has no impact on the (ergodic) primary performance.
			For RIScatter, different state combinations correspond to different messages and the impact reduces to dynamic passive beamforming upon successful backscatter decoding.
			% However, different combinations correspond to different backscatter messages and RIScatter impact is determined dynamic passive beamforming once successfully decoded.
		\end{remark}

@@

			% \resizebox{0.8\columnwidth}{!}{
			% 	\input{assets/illustration/energy_distribution.tex}
			% }
			\includegraphics[width=0.8\columnwidth]{assets/illustration/energy_distribution.jpg}

@@

	Problem \eqref{op:weighted_sum_rate} generalizes conventional \gls{bc} by allowing \gls{csi}- and \gls{qos}-adaptive input distribution and decision region design.
	It also generalizes the discrete \gls{ris} phase shift selection problem by stochastic reflection that relaxes the feasible domain from the vertices of $M$-dimensional probability simplex to the simplex itself.
	Problem \eqref{op:weighted_sum_rate} is highly non-convex and we propose a \gls{bcd} algorithm that iteratively updates $\{\boldsymbol{p}_k\}_{k \in \mathcal{K}}$, $\boldsymbol{w}$ and $\boldsymbol{t}$.

@@

		\begin{remark}
			Energy detector design has no impact on the primary achievable rate since the composite channel \eqref{eq:equivalent_channel} can always be reconstructed after backscatter decoding and re-encoding.
			It suggests any thresholding scheme maximizing the total backscatter rate \eqref{eq:backscatter_mutual_information} is also optimal for problem \eqref{op:decision_threshold}.
			\label{re:backscatter_decision}
		\end{remark}

@@

\begin{section}{Simulation Results}
	In this section, we provide numerical results to evaluate the proposed input distribution, active beamforming, and backscatter decision designs for the considered RIScatter system.
	We assume the AP-user distance is \SI{10}{\meter} and at least one RIScatter nodes are randomly dropped in a disk centered at the user with radius $r$.
	% The \gls{ap} is with an average transmit power budget $P=\SI{36}{\dBm}$, all nodes employs $M$-\gls{qam} with amplitude scattering ratio $\alpha=0.5$, and the user is with average noise power $\sigma_v^2=\SI{-40}{\dBm}$.
	The \gls{ap} is with an average transmit power budget $P=\SI{36}{\dBm}$ and all nodes employs $M$-\gls{qam} with $\alpha=0.5$.
	For all channels involved, we consider a distance-dependent path loss model
	\begin{equation}
		L(d) = L_0 \biggl(\frac{d_0}{d}\biggr)^\gamma,
	\end{equation}
	together with a Rician fading model
	\begin{equation}
		\boldsymbol{H} = \sqrt{\frac{\kappa}{1+\kappa}} \bar{\boldsymbol{H}} + \sqrt{\frac{1}{1+\kappa}} \tilde{\boldsymbol{H}},
	\end{equation}
	where $d$ is the transmission distance, $L_0=-\SI{30}{\dB}$ is the reference path loss at $d_0=\SI{1}{\meter}$, $\kappa$ is the Rician K-factor, $\bar{\boldsymbol{H}}$ is the deterministic line-of-sight component with unit-magnitude entries, and $\tilde{\boldsymbol{H}}$ is the Rayleigh fading component with standard \gls{iid} \gls{cscg} entries.
	We choose $\gamma_{\text{D}}=2.6$, $\gamma_{\text{F}}=2.4$, $\gamma_{\text{B}}=2$, and $\kappa_{\text{D}}=\kappa_{\text{F}}=\kappa_{\text{B}}=5$ for direct, forward and backward links.
	The finite decision threshold domain $\mathcal{T}$ is obtained by $b$-bit uniform discretization over the critical interval defined by the confidence bounds of edge hypotheses (i.e., lower bound of $\mathcal{H}_1$ and upper bound of $\mathcal{H}_L$) with confidence $1-\varepsilon$.
	We set $b=9$ and $\varepsilon={e-3}$.
	All achievable rate points/regions are averaged over $10^3$ channel realizations.

@@

		\begin{subsubsection}{Convergence}
			\begin{figure}[!t]
				\centering
				\resizebox{0.75\columnwidth}{!}{
					\input{assets/simulation/wsr_convergence.tex}
				}
				\caption{Typical convergence curves at $\rho=0$ for $Q=4$, $K=8$, $M=2$, $N=20$, $\sigma_v^2=\SI{-40}{\dBm}$ and $r=\SI{2}{\meter}$.}
				\label{fg:wsr_convergence}
			\end{figure}

			In Fig.~\ref{fg:wsr_convergence}, we plot the weighted sum of primary and total backscatter rates at $\rho=0$ for \gls{kkt}, \gls{pgd} and \gls{bcd} algorithms on the first call.
			For $K=8$ and $M=2$, Algorithm \ref{al:input_distribution} typically takes around {100} fast iterations to converge to the \gls{kkt} input distribution.
			For $Q=4$, around {10} iterations are required for Algorithm \ref{al:active_beamforming} to converge.
			Overall, the initial \gls{bcd} algorithm requires at most {5} iterations to converge.
			At the following points (not presented here), the convergence of all three algorithms are much faster thanks to the progressive initialization.
			We conclude the proposed algorithms have good convergence performances.
		\end{subsubsection}
	\end{subsection}

@@

			\caption{Typical achievable rate region/points of scattering applications for $Q=1$, $K=1$, $M=4$, $N=10^3$, $\sigma_v^2=\SI{-40}{\dBm}$ and $r=\SI{2}{\meter}$.}
